### 2017-11-15 Draft2/Indexing.Dynamic.tex
Prune indexing section

The following algorithm implements $\FuncGetItem$ for dynamic arrays. (A $\FuncSetItem$ function can be implemented in a similar fashion.)

{\HdrNote} My algorithms do not validate any arguments; all arguments passed to them are assumed to be valid.

\begin{algorithm}
	\begin{algorithmic}[1]
		\Function{$\FuncGetItem$}{$\VarList,\ \ParamIndex$}
			\State \Return $\VarList.\FieldBuffer[\ParamIndex]$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

This algorithm runs in $O(1)$ time. It also has the benefit of using very few instructions, which is a benefit that will not be shared with growth arrays.

>>>>

The algorithm for indexing a dynamic array is trivial, and will not be shown here. It simply reads or writes to the $i$th element of $\VarList.\FieldBuffer$ (where $i$ is zero-based). It is well-known that this algorithm runs in $O(1)$ time. It also has the benefit of using very few instructions, which will not be the case for growth arrays' algorithm.

### 2017-11-15 Draft2/Indexing.Growth.tex
Exterminate more todos
Exterminate things
Add pictures, [WIP] title all the algos
Prune indexing section

Despite growth arrays' relatively poor indexing performance, I include this section for two reasons. 1) Dynamic arrays are indexed frequently. In order for growth arrays to be a viable replacement for them, it should be possible to index growth arrays too. 2) I wish to show that it is possible to index growth arrays in constant time.

@@

The first algorithm I will demonstrate is a na\"{i}ve implementation that runs in logarithmic time:

\begin{algorithm}[H]
	\begin{algorithmic}
		\Function{$\FuncGetItem$}{$\VarList,\ \ParamIndex$}
			\State $\LclIndex \gets \ParamIndex$
			\For{$\LclBuffer \TextIn \VarList.\FieldTail$}
				\If{$\LclIndex < \LclBuffer.\FieldLength$}
					\State \Return $\LclBuffer[\LclIndex]$
				\EndIf
				\State $\LclIndex \gets \LclIndex - \LclBuffer.\FieldLength$
			\EndFor
			\State \Return $\VarList.\FieldHead[\LclIndex]$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

I know that this algorithm runs in $O(\log n)$ time since it was shown earlier that the size of the tail, $\VarNumPointersTail$, equals $\VarUseful$, and (from Lemma \ref{lem:VarUsefulIsOLogN}) that $\VarUseful = O(\log n)$. Aside from iteration over the tail, all statements run in constant time, so the time complexity is precisely $O(\log n)$.

I present this algorithm alongside the constant time one because it is easier to understand, and the latter is often slower if special circumstances are not met.

@@

Suppose I want to get the $i$th element of a growth array, where $i$ is zero-based. I assume that $i$ is a valid index, or that $0 \leq i < \VarList.\FieldSize$. In order to locate the desired element, I must find two things: the buffer that holds the item, and the index of the item within that buffer.

@@

\begin{algorithm}[H]
	\begin{algorithmic}
		\Function{$\FuncGetBuffer$}{$\VarList,\ \LclIndexBuffer$}
			\If{$\LclIndexBuffer < \VarList.\FieldTail.\FieldSize$}
				\State \Return $\VarList.\FieldTail[\LclIndexBuffer]$
			\Else
				\State \Return $\VarList.\FieldHead$
			\EndIf
		\EndFunction
	\end{algorithmic}
\end{algorithm}

@@

If I define a helper function $\FuncDecompose$ that returns the $\VarIndexBuffer$ and $\VarIndexElement$ associated with $i$ in an ordered pair, then $\FuncGetItem$ may be written as follows:

\begin{algorithm}[H]
	\begin{algorithmic}
		\Function{$\FuncGetItem$}{$\VarList,\ \ParamIndex$}
			\State $(\LclIndexBuffer,\ \LclIndexElement) \gets \FuncDecompose(\ParamIndex)$
			\State \Return $\VarList.\FuncGetBuffer(\LclIndexBuffer)[\LclIndexElement]$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

Now, the task is to find formulae for $\VarIndexBuffer$ and $\VarIndexElement$ in terms of $i$, in order to implement $\FuncDecompose$.

% Todo: There's an awful lot of lemmas. Make some into theorems.
\begin{lemma}
	The formulae for $\VarIndexBuffer$ and $\VarIndexElement$ are
	\begin{align*}
	\VarIndexBuffer &= \VarUseful|_{n = i + 1}\\
	\VarIndexElement &= i - \VarGrowSeq_{\VarIndexBuffer - 1}
	\end{align*}
	(When $\VarIndexBuffer = 0$, $\VarGrowSeq_{-1} = 0$ by convention.)
\end{lemma}

% Todo: "As shown earlier"
\begin{proof}
	Appending new items does not change the index of an item that is already in the list. Thus, finding the element at index $i$ is equivalent to finding the last element when $n = i + 1$.

@@

	To determine $\VarIndexElement$, consider the identity:
	\begin{align*}
	n = \text{\# of items in tail} + \text{\# of items in head}
	\end{align*}
	First, I will determine the number of items in the tail. In the case where $\VarIndexBuffer = 0$, then because $\VarNumPointersTail = \VarIndexBuffer$, the tail contains $0$ buffers and thus $0$ items (which equals $\VarGrowSeq_{-1}$). If $\VarIndexBuffer > 0$, the number of items in the tail is the last size at which $\FuncGrow$ was called, or the last term of $\VarGrowSeq$. This quantity is $\VarGrowSeq_{\VarUseful - 1}$, or $\VarGrowSeq_{\VarIndexBuffer - 1}$.

@@

Using the formulae for $\VarUseful$ and $\VarGrowSeq_i$, I now implement the $\FuncDecompose$ function:

\begin{algorithm}[H]
	\begin{algorithmic}
		\Function{$\FuncDecompose$}{$\ParamIndex$}
			\State $\LclIndexBuffer \gets \max(\left\lceil \log_{\VarGrowthFactor} (\ParamIndex + 1) - \log_{\VarGrowthFactor} \VarInitCapacity \right\rceil, 0)$
			\If{$\VarIndexBuffer > 0$}
				\State $\LclIndexElement \gets \ParamIndex - \VarGrowthFactor^{\LclIndexBuffer - 1} \times \VarInitCapacity$
			\Else
				\State $\VarIndexElement \gets \ParamIndex$
			\EndIf
			\State \Return $(\LclIndexBuffer,\ \LclIndexElement)$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

>>>>

Although growth arrays have poor random access performance, I include this section for two reasons. 1) Dynamic arrays are indexed frequently. In order for growth arrays to be a viable replacement for them, it should be possible to index growth arrays too. 2) I wish to show that it is possible to index growth arrays in constant time.

@@

The first algorithm I will demonstrate is a na\"{i}ve implementation that runs in logarithmic time.

{\HdrNote} My algorithms assume all arguments passed to them are valid. Thus, arguments are never checked.

\begin{algorithm}[H]
	\caption{Random access \TextGrowthArray, logarithmic time}
	\begin{algorithmic}
		\Function{$\FuncGetItem$}{$\VarList,\ \ParamIndex$}
			\State $\LclIndex \gets \ParamIndex$
			\For{$\LclBuffer \TextIn \VarList.\FieldTail$}
				\If{$\LclIndex < \LclBuffer.\FieldLength$}
					\State \Return $\LclBuffer[\LclIndex]$
				\EndIf
				\State $\LclIndex \gets \LclIndex - \LclBuffer.\FieldLength$
			\EndFor
			\State \Return $\VarList.\FieldHead[\LclIndex]$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

I know that this algorithm runs in $O(\log n)$ time since it was shown earlier that the size of the tail, $\VarNumPointersTail$, equals $\VarUseful$, and Lemma \ref{lem:VarUsefulIsOLogN} states that $\VarUseful = O(\log n)$. Aside from iteration over the tail, all statements run in constant time, so the time complexity is precisely $O(\log n)$.

I present this algorithm alongside the constant time one since it is easier to understand, and the latter is often slower if special circumstances are not met.

@@

Suppose I want to get the $i$th element of a growth array, where $i$ is zero-based. I assume that $i$ is a valid index, or that $0 \leq i < \VarList.\FieldSize$. In order to locate the desired element, I must find two things: the buffer that holds the item, and the index of the item within that buffer. For example, referring to Figure \ref{fig:1}: if $i$ was $10$, I would want the item at index $2$ of the tail's third buffer. If $i$ was $20$, I would want the item at index $4$ of the head.

@@

\begin{algorithm}[H]
	\caption{\TextHelperFunction}
	\begin{algorithmic}
		\Function{$\FuncGetBuffer$}{$\VarList,\ \LclIndexBuffer$}
			\If{$\LclIndexBuffer < \VarList.\FieldTail.\FieldSize$}
				\State \Return $\VarList.\FieldTail[\LclIndexBuffer]$
			\Else
				\State \Return $\VarList.\FieldHead$
			\EndIf
		\EndFunction
	\end{algorithmic}
\end{algorithm}

@@

If I define a helper function $\FuncDecompose$ that splits $i$ into the ordered pair $(\VarIndexBuffer,\ \VarIndexElement)$, then $\FuncGetItem$ may be written as follows:

\begin{algorithm}[H]
	\caption{Random access \TextGrowthArray, constant time}
	\begin{algorithmic}
		\Function{$\FuncGetItem$}{$\VarList,\ \ParamIndex$}
			\State $(\LclIndexBuffer,\ \LclIndexElement) \gets \FuncDecompose(\ParamIndex)$
			\State \Return $\VarList.\FuncGetBuffer(\LclIndexBuffer)[\LclIndexElement]$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

Now, the task is to find formulae for $\VarIndexBuffer$ and $\VarIndexElement$ in terms of $i$, in order to implement $\FuncDecompose$. Referring to Figure \ref{fig:1}: $\FuncDecompose(10)$ should return $(2,\ 2)$. $\FuncDecompose(20)$ should return $(3,\ 4)$.

\begin{lemma}
	The formulae for $\VarIndexBuffer$ and $\VarIndexElement$ are
	\begin{align*}
	\VarIndexBuffer &= \VarUseful|_{n = i + 1}\\
	\VarIndexElement &= i - \VarGrowSeq_{\VarIndexBuffer - 1}
	\end{align*}
	(By convention, $\VarGrowSeq_{-1} = 0$.)
\end{lemma}

\begin{proof}
	Appending new items does not change the index of an item that is already in the list. Thus, finding the element at index $i$ is equivalent to finding the last element when $n = i + 1$.

@@

	To determine $\VarIndexElement$, consider the identity:
	\begin{align*}
	n = \text{\# of items in tail} + \text{\# of items in head}
	\end{align*}
	({\HdrNote} The first quantity on the right-hand side is not $\VarNumPointersTail$. $\VarNumPointersTail$ is the number of buffer pointers in the tail.)

	First, I will determine how many items the tail holds. In the case where $\VarIndexBuffer = 0$, then because $\VarNumPointersTail = \VarIndexBuffer$, the tail contains $0$ buffer pointers and thus $0$ items (which equals $\VarGrowSeq_{-1}$). If $\VarIndexBuffer > 0$, the number of items in the tail is the last size at which $\FuncGrow$ was called, or the last term of $\VarGrowSeq$. This quantity is $\VarGrowSeq_{\VarUseful - 1}$, or $\VarGrowSeq_{\VarIndexBuffer - 1}$.

@@

Using the formulae $\VarUseful = \max(\ExprMessy, 0)$ and $\VarGrowSeq_k = \VarGrowthFactor^k \VarInitCapacity$ for $k \geq 0$, I now implement the $\FuncDecompose$ function:

\begin{algorithm}[H]
	\caption{\TextHelperFunction}
	\begin{algorithmic}
		\Function{$\FuncDecompose$}{$\ParamIndex$}
			\State $\LclIndexBuffer \gets \max(\left\lceil \log_{\VarGrowthFactor} (\ParamIndex + 1) - \log_{\VarGrowthFactor} \VarInitCapacity \right\rceil, 0)$
			\If{$\VarIndexBuffer > 0$}
				\State $\LclIndexElement \gets \ParamIndex - \VarGrowthFactor^{\LclIndexBuffer - 1} \times \VarInitCapacity$
			\Else
				\State $\VarIndexElement \gets \ParamIndex$
			\EndIf
			\State \Return $(\LclIndexBuffer,\ \LclIndexElement)$
		\EndFunction
	\end{algorithmic}
\end{algorithm}
